Quantifying the ease of viral escape from broad and narrow antibodies to influenza hemagglutinin

### 2017-10-19 paper/manuscript.tex

Specifically, for S139/1, we observed strong escape mutations at sites 156, 158, and 193 (Figure~\ref{fig:structures}A; sites are in H3 numbering). 
These three sites surround HA's receptor-binding pocket, and fall directly in the physical binding footprint of the antibody~\citep{lee2012heterosubtypic}.
These are also the same three sites where \citet{yoshida2009cross} selected escape mutants in H1, H2, and H3 HAs. 
Our data show that numerous different amino-acid mutations at each site confer neutralization resistance.
The mutation with the largest effect, D158N, introduces an N-linked glycosylation site.

The mutation contributing the greatest fraction of virions surviving selection, G158N, introduces a glycosylation motif (N-X-S/T) at this site.
C179 and FI6v3, which share very similar binding footprints and angles of approach~\citep{dreyfus2013structure}, select for identical escape mutations at several sites (Figure~\ref{fig:structures}B). 
Both antibodies select for the K280S, K280T, and N291S mutations, all of which introduce glycosylation motifs in close proximity to the epitope and all have similar effect sizes in conferring escape from either antibody.

Despite the similarities in the epitope-targeting of C179 and FI6v3, we also see selection for distinct escape mutations with either antibody.
The mutation with the strongest effect size from selection by C179 occurs at site 38, while we do not see observable escape mutations at this site by selection with FI6v3.
Interestingly, we found that HA subtypes susceptible to binding or neutralization by C179 tend to possess a histidine at this site, but subtypes not bound or neutralized by C179 possess a glutamine or serine (Figure~\ref{fig:structures}C). 
This would suggest that susceptibility to neutralization by C179 may in part be determined by the amino-acid identity at site 38.
Furthermore, the difference in breadth between C179 and FI6v3 has been attributed to FI6v3 being able to accommodate a glycan on the Asn at site 38 which is present in group 2 but not group 1 HA's, while group 1-specific stalk-targeting antibodies clash with this glycan~\citep{corti2011neutralizing, sui2009structural, ekiert2009antibody}. 
The H38S mutation selected by C179 does not introduce a glycosylation motif, which indicates that introduction of a glycan at this site is not the only means of mediating escape from this group 1-specific anti-stalk antibody.
The amino-acid identity at site 111 of HA2 can change the orientation of a conserved Trp21 in HA2, also resulting in group 1 and group 2 differences in the binding ability of stalk-targeting antibodies. An H111T mutation in the A/South Carolina/1/1918 (H1N1) has been reported to abrogate C179 binding~\citep{dreyfus2013structure}.
However, we did not observe escape mutations at site 111, but this may be due to strain-specific differences.

>>>>

Specifically, for broad anti-receptor binding pocket antibody S139/1, we observe strong escape mutations at sites 156, 158, and 193 (Figure~\ref{fig:structures}A; sites are in H3 numbering). 
These three sites fall directly in the physical binding footprint of the antibody~\citep{lee2012heterosubtypic}.
These are also the same three sites where \citet{yoshida2009cross} selected escape mutants in H1, H2, and H3 HAs. 
Our data show that numerous different amino-acid mutations at each site confer neutralization resistance.
The mutation with the largest effect, D158N, introduces an N-linked glycosylation motif.

Although anti-stalk antibodies C179 and FI6v3 only select mutations with small effects, these mutations almost all fall in the physical binding footprints of the antibodies (Figure~\ref{fig:structures}A).
The two antibodies have similar epitopes and angles of approach~\citep{dreyfus2013structure}, and they select identical mutations at several sites (Figure~\ref{fig:structures}B). 
The three largest-effect mutations for FI6v3 (K280S, K280T, and N291S) all introduce N-linked glycosylation motifs near the epitope, and all three mutations also have a similar effect size for C179.

However, C179 also selects several mutations that do not have any apparent effect on recognition by FI6v3 (Figure~\ref{fig:structures}A, Figure~\ref{suppfig:FI6v3logo}).
The most notable of these C179-specific mutations occurs at site 38.
The additional breadth of FI6v3 over antibodies such as C179 that neutralize only group 1 HAs has been attributed to FI6v3's ability to accommodate a glycan on the asparagine at site 38 that is present in group 2 but not group 1 HAs~\citep{corti2011neutralizing,sui2009structural,ekiert2009antibody}. 
However, the H38S mutation that has the largest effect on C179 resistance in our experiments does not introduce a glycosylation motif, showing that there are also other ways to escape anti-stalk antibodies at this site.
Interestingly, group 1 HA subtypes that are susceptible by C179 tend to possess a histidine at site 38, but subtypes that are not bound or neutralized by C179 often possess a serine (Figure~\ref{fig:structures}C). 
%The amino-acid identity at site 111 of HA2 can change the orientation of a conserved Trp21 in HA2, also resulting in group 1 and group 2 differences in the binding ability of stalk-targeting antibodies. An H111T mutation in the A/South Carolina/1/1918 (H1N1) has been reported to abrogate C179 binding~\citep{dreyfus2013structure}.
%However, we did not observe escape mutations at site 111, but this may be due to strain-specific differences.

\comment{still need to add something about signal peptide mutation, maybe refer to H17* antibodies too}
